%---------------------------------------------------------------------------%
%->> Frontmatter
%---------------------------------------------------------------------------%
%-
%-> 生成封面
%-
\maketitle% 生成中文封面
\MAKETITLE% 生成英文封面
%-
%-> 作者声明
%-
\makedeclaration% 生成声明页
%-
%-> 中文摘要
%-
\intobmk\chapter*{摘\quad 要}% 显示在书签但不显示在目录
\setcounter{page}{1}% 开始页码
\pagenumbering{Roman}% 页码符号

面向国产处理器平台完善 Android 应用执行环境，是推进基础软件生态建设的重要环节。AOSP 15
官方主线尚未提供面向 LoongArch 架构的 ART 执行引擎实现，这使得 Android 应用在该平台上的
稳定运行与性能优化缺少关键运行时基础。本文以 AOSP 15 的 Android Runtime（ART）为研究
对象，围绕 LoongArch 平台执行引擎的适配设计、正确性验证与性能优化展开系统研究。

本文首先面向 LoongArch 平台完成了 ART 执行引擎关键模块的适配，系统补齐了汇编器、C++
解释器、Nterp、JIT 编译器后端、JNI 编译器、运行时快速路径以及部分 Intrinsic 支持，构建了
LoongArch 平台上可运行的 ART 执行环境。其次，本文建立了覆盖编译、部署、target 回归、
libcore 测试与 benchmark 评测的统一验证链路，形成了固定 JIT baseline、日志归档与热点分析
方法，为移植结果和后续优化实验提供了可复现支撑。

在性能优化方面，本文基于统一 baseline 对典型 workload 的热点行为进行分析，围绕
\texttt{DaCapo/pmd}、\texttt{SPECjvm2008/mpegaudio} 与 Scimark 数值子项提出并验证了
三项正式优化方案：JIT 分层热度阈值调整方案、运行时字符串搬运批量化方案，以及 LoongArch64
LSX 浮点 SIMD 向量化方案。实验结果表明，JIT 分层热度阈值调整方案在 \texttt{pmd} 上存在
可复现的正收益样本，但重复测量仍显示较大方差；运行时字符串搬运批量化方案仅呈现弱正收益，
更适合作为字符串密集场景的定向微优化；LoongArch64 LSX 浮点 SIMD 向量化方案则在
\texttt{SPECjvm2008/scimark.lu.small} 上取得了稳定且显著的性能提升，是当前证据最完整的
主收益结论。进一步结合方案叠加负例与若干失败方案的对照分析，本文归纳出 LoongArch 平台 ART
执行引擎优化应优先命中主导热点、严格控制全局副作用，并明确不同 workload 的收益边界。

研究结果表明，AOSP 15 ART 在 LoongArch 平台上的执行引擎适配是可行的；在统一验证链路与固定
baseline 约束下，围绕执行引擎热点开展定向优化能够获得可复现收益。本文工作为国产处理器平台上
Android 运行时的进一步完善与性能提升提供了工程基础和方法参考。

\keywords{LoongArch，Android Runtime，执行引擎适配，JIT 编译，性能优化}% 中文关键词
%-
%-> 英文摘要
%-
\intobmk\chapter*{Abstract}% 显示在书签但不显示在目录

Improving the Android application runtime environment on domestic processor
platforms is an important part of building a complete system software ecosystem.
However, AOSP 15 does not provide an official ART execution engine
implementation for the LoongArch architecture, which leaves a critical gap in
application execution support on this platform. This thesis takes Android
Runtime (ART) in AOSP 15 as the research target and studies execution engine
adaptation, correctness validation, and performance optimization for
LoongArch.

First, this work completes the adaptation of key ART execution engine
components for LoongArch, including the assembler, C++ interpreter, Nterp, the
JIT compiler backend, the JNI compiler, runtime fast paths, and selected
intrinsics, thereby establishing a runnable ART execution environment on the
target platform. Second, it builds a unified validation workflow covering
compilation, deployment, target-side regression tests, libcore tests, and
benchmark evaluation, and further establishes a fixed JIT baseline together
with reproducible logging and hotspot analysis procedures.

For performance optimization, this thesis analyzes hotspot behavior under a
fixed baseline and studies three formal schemes on representative workloads: a
tiered JIT hotness threshold adjustment, a runtime string bulk-copy
optimization, and a LoongArch64 LSX floating-point SIMD vectorization scheme.
The results show that the tiered JIT threshold adjustment has
reproducible positive samples on \texttt{DaCapo/pmd} but still exhibits notable
variance under repeated measurements; the runtime string bulk-copy optimization
only delivers weak gains and is therefore better characterized as a targeted
micro-optimization for string-intensive workloads; and the LoongArch64 LSX
floating-point SIMD scheme achieves a stable and substantial improvement on
\texttt{scimark.lu.small}, making it the strongest positive result supported by
the current evidence. By further comparing stacking conflicts and several
negative results, this thesis concludes that optimization of the ART execution
engine on LoongArch should prioritize dominant hotspots, strictly control
global side effects, and explicitly state workload applicability boundaries.

The results demonstrate that adapting the AOSP 15 ART execution engine to the
LoongArch platform is feasible, and that targeted hotspot-driven optimizations
under a unified validation workflow and a fixed baseline can produce
reproducible benefits. This work provides both an engineering foundation and a
practical methodology reference for further improving Android runtime support on
domestic processor platforms.

\KEYWORDS{LoongArch, Android Runtime, execution engine adaptation, JIT compilation, performance optimization}% 英文关键词
%---------------------------------------------------------------------------%
